\documentclass[compress]{beamer}

\mode<presentation>
\usetheme{Madrid}
\definecolor{Columbia}{RGB}{185,217,235}
\definecolor{Columbia2}{RGB}{0,51,160}
\definecolor{Columbia3}{RGB}{0,114,206}
\setbeamercolor{title}{fg=Columbia2}
\setbeamercolor{frametitle}{fg=Columbia2}
\setbeamercolor{block title}{bg=Columbia, fg=Columbia2}
%\setbeamercolor{block body}{fg=Columbia2}
\setbeamercolor{structure}{fg=Columbia}
\setbeamercolor{item projected}{fg=white}
\setbeamercolor{item}{fg=Columbia2}
\setbeamercolor{subitem}{fg=Columbia2}
\setbeamercolor{section in toc}{fg=Columbia2}
\setbeamercolor{description item}{fg=Columbia}
\setbeamercolor{caption name}{fg=Columbia2}
\setbeamercolor{button}{fg=Columbia2}
\usepackage{graphics}
\usepackage{geometry}
\usepackage{booktabs}
\usepackage{tikz}
\usepackage{amsmath}
\usepackage{bbm}
\usetikzlibrary{decorations.pathreplacing}
\usepackage{multirow, makecell}
\usepackage{float}
\usepackage{fancyvrb}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{adjustbox}
\usepackage{threeparttable}
\usepackage{hyperref}
\usepackage[scaled=0.92]{helvet}
\newenvironment{wideitemize}{\itemize\addtolength{\itemsep}{10pt}}{\enditemize}
\newenvironment{wideenumerate}{\enumerate\addtolength{\itemsep}{10pt}}{\endenumerate}
%\hypersetup{
%colorlinks=true,
%linkcolor=black,
%filecolor=green, 
%urlcolor=blue,
%}
\beamertemplatenavigationsymbolsempty
\setbeamercolor{author in head/foot}{bg=Columbia2, fg=white}
\setbeamercolor{title in head/foot}{bg=Columbia3, fg=white}
\setbeamercolor{date in head/foot}{bg=Columbia, fg=Columbia2}
\setbeamercolor{section in head/foot}{bg=Columbia, fg=Columbia2}
\setbeamercolor{headline}{bg=Columbia}
\setbeamertemplate{footline}{
    \leavevmode%
    \hbox{%
        \begin{beamercolorbox}[wd=.333333\paperwidth,ht=2.25ex,dp=1ex,center]{author in head/foot}%
            \usebeamerfont{author in head/foot}\insertshortauthor
        \end{beamercolorbox}%
        \begin{beamercolorbox}[wd=.333333\paperwidth,ht=2.25ex,dp=1ex,center]{title in head/foot}%
            \usebeamerfont{title in head/foot}\insertshorttitle
        \end{beamercolorbox}%
        \begin{beamercolorbox}[wd=.333333\paperwidth,ht=2.25ex,dp=1ex,right]{date in head/foot}%
            \usebeamerfont{date in head/foot}\insertshortdate{}\hspace*{2em}
            \insertframenumber{} / \inserttotalframenumber\hspace*{2ex} 
        \end{beamercolorbox}}%
        \vskip0pt%
    }
\setbeamercolor{page number in head/foot}{fg=black}
\setbeamertemplate{section in toc}[sections numbered]
\setbeamertemplate{subsection in toc}{\leavevmode\leftskip=3em\rlap{\hskip-1.75em\inserttocsectionnumber.\inserttocsubsectionnumber}\inserttocsubsection\par}
\setbeamerfont{subsection in toc}{size=\footnotesize}
%\setbeamertemplate{headline}{%
  %\begin{beamercolorbox}[ht=5.5ex]{section in head/foot}
    %\vskip2pt\insertnavigation{0.33\paperwidth}\vskip2pt
  %\end{beamercolorbox}%
%}


\makeatletter
\let\@@magyar@captionfix\relax
\makeatother


\title[Recitation 7]{Recitation 7} % Change this regularly
\author[Seung-hun Lee]{Seung-hun Lee}
\institute[Columbia University]{Columbia University}

\date[]{}

\begin{document}
\begin{frame}
\titlepage
\end{frame}

%%%%%%%%%%%%% Section 1. 


% Capture in one slide
% Separate slide for questions
\begin{frame}
\frametitle{Panel regression}
Motivation
\begin{itemize}
\item \textbf{Panel data}: We observe multiple individuals for multiple periods of time.
\[
Y_{it} = \beta_0 + \beta_1X_{1,it}+ ... +\beta_kX_{k,it}+u_{it}
\]
$i=1,2,...,N \to$  individuals, $t=1,2,...,T\to$  time periods.
\item Balanced: There are $T$ datasets for each of the $N$ individuals.
\item Unbalalced: There are $t\leq T$ datapoints for some of the $N$ individuals.
\end{itemize}

Advantages
\begin{itemize}
\item Panel data allows us to use more datasets. 
\item Panel data allows us to control for \textbf{unobserved heterogeneity} that are
\begin{enumerate}
\item  different accross $N$ entities but always remain same for $T$ periods in a given state (\textbf{cross section fixed effect})
\item different accross $T$ times periods but remains the same for all $N$ entities in a particular time period  (\textbf{time fixed effects})
\item both of 1) and 2). (\textbf{two-way fixed effects})
\end{enumerate}
\end{itemize}
\end{frame}


\begin{frame}
\frametitle{Panel regression}
Key differences
\begin{itemize}
\item Suppose that $T=2$ and we are interested in the relationship between vehicle related fatality rate (deaths per 10,000 people) and the beer tax. Suppose that we get these result for the two years
\[
\begin{aligned}
\hat{Y}_{i1} =&2.01 +&0.15X_{i1}\\
                    &(0.15)&(0.20) \\
\hat{Y}_{i2} =&1.86 +&0.44X_{i2}\\
                    &(0.11)&(0.20) \\                    
\end{aligned}
\]
\item  In such case, one might suspect that there is an omitted variable bias that affects these coefficients.
\begin{itemize}
\item Omitted variable specific to the states (Strictness of the relevant law)
\item Time-trends? (Specific to each of years 1 and 2)
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Panel regression}
How can panel regression do better?
\begin{itemize}
\item Let $Z_i$ denote the strictness of state laws on DUI that are unchanging. 
\item Now write
\[
\begin{aligned}
Y_{i1}& = \beta_0 + \beta_1X_{i1}+\beta_2 Z_{i}+u_{i1} \\                
Y_{i2}& = \beta_0 + \beta_1X_{i2}+\beta_2 Z_{i}+u_{i2} \\                
\end{aligned}
\]
\item Subtract the second equation from the first to get
\[
(Y_{i2}-Y_{i1}) = \beta_1(X_{i2}-X_{i1}) +\beta_2(Z_{i}-Z_{i}) + u_{i2}-u_{i1}
\]
With $Z_i$ being the same for all periods, the above equation is reduced to
\[
(Y_{i2}-Y_{i1}) = \beta_1(X_{i2}-X_{i1}) +(u_{i2}-u_{i1})
\]
\item The $Z_i$ variable has no role in this equation - because it is now gone.
\item If we estimate this particular $\beta_1$, we can obtain much more accurate estimates of the effect of bear tax on fatality rate. 
\end{itemize}
\end{frame}


\begin{frame}
\frametitle{Panel regression}
Specific methodologies (cross-sectional FE)
\begin{itemize}
\item There are two ways of estimating the data when $T\geq 3$
\item Least square dummy variables: Include $N-1$ individual dummies
\item Within estimation: Subtract ``demeaned" equation from the original
\item Use: 
\[
Y_{it}=\beta_0 + \beta_1X_{it}+\beta_2 Z_{i}+u_{it} \tag{1}
\]
where $Z_i$ is the cross section fixed effect.
\item Define  $\alpha_i = \beta_0 + \beta_1Z_i$. Then the above equation can be written as
\[
Y_{it}=\beta_1X_{it}+\alpha_{i}+u_{it} \tag{2}
\]
\item $\alpha_i$ term can be thought of as an effect of being an entity $i$, which is \textbf{correlated with} $X_{it}$
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Panel regression}
LSDV method
\begin{itemize}
\item Define a new variable $D_{ki}$ as follows
\[
D_{ki} = \begin{cases} 1 & \text{If $i=k$} \\
                                     0 & \text{Otherwise } \end{cases} , \ k\in\{1,2,...,N\}
\] 
\item Since we are going to include $\beta_0$, a common intercept, in our regression we need to remove one of the $N$ (dummy variable trap)
\item Then we can write
\[
Y_{it} = \beta_0 +\beta_1X_{it}+\delta_2D_{2i} + ... + \delta_ND_{Ni}+u_{it} \ \tag{LSDV}
\]
\item This equation gives different intercepts for each $i$ (can you see why?), while keeping the slope on $X_{it}$ constant at $\beta_1$ 
\item Control for unobserved cross section fixed effect by allowing the intercept to differ by each $i$
\end{itemize}
\end{frame}


\begin{frame}
\frametitle{Panel regression}
WE method
\begin{itemize}
\item Define $\bar{X}_i, \bar{Y}_i$ as sample mean of $X_{it}, Y_{it}$ for given $i$ over all possible $t$'s. 
\[
\bar{X}_i = \frac{1}{T}\sum_{t=1}^TX_{it}
\]
Consequently, $\bar{Y}_i$ can be written as
\[
\bar{Y}_i = \frac{1}{T}\sum_{t=1}^TY_{it}=\frac{1}{T}\sum_{t=1}^T\left(\beta_1 X_{it} +\alpha_i +u_{it}\right)=\beta_1 \bar{X}_i +\alpha_i + \bar{u}_{i}
\]
\item Subtract $Y_{it}$ by $\bar{Y}_i$ to get
\[
\begin{aligned}
Y_{it}-\bar{Y}_i &= \beta_1(X_{it}-\bar{X}_i) + (u_{it}-\bar{u}_i) \implies \tilde{Y}_{it}= \beta_1 \tilde{X}_{it}+\tilde{u}_{it}\\
\end{aligned}
\]
\item This process gets rid of $\alpha_i$. Then, apply OLS estimation on this equation to get the within estimator
\end{itemize}
\end{frame}


\begin{frame}
\frametitle{Panel regression}
TWFE methods
\begin{itemize}
\item We have a DGP
\[
Y_{it}=\beta_1 X_{it} + \alpha_i +\lambda_t + u_{it}
\]
\item LSDV: With an overall constant $\beta_0$, we can put $N-1$ individual and $T-1$ time dummies
\item WE: Demeaning should be done in the following method
\[
Y_{it}-\bar{Y}_i -\bar{Y}_t +\bar{Y}
\]
This (and only this) would allow us to get rid of both the $\alpha_i$ individual fixed effect and the $\lambda_t$ time effects
\end{itemize}
\end{frame}


\begin{frame}
\frametitle{Panel regression}
Least square assumptions for panels
\begin{itemize}
\item  [\textbf{P1}]: $E[u_{it}|X_{i1},..,X_{iT},\alpha_i]=0$. It means that the conditional mean of the $u_{it}$ term does not depend on any of the $X_{it}$ values for entity $i$, whether in the future or in the past. 
\item [\textbf{P2}]: $(X_{i1},..,X_{iT},u_{i1},...u_{iT})$ is IID across $i=1,..,n$. \textbf{This does not rule out the correlation between $u_{it},u_{ij}$ within entity $i$ for different $j$ and $t$}, allowing serial correlation within the same entity
 \item[\textbf{P3}]: $(X_{it},u_{it})$ have nonzero finite fourth moments (outliers are very unlikely) so that the panel estimators have a distribution
\item [\textbf{P4}]: There is no perfect multicollinearity
\end{itemize}
$\to$ Because of P2, we need to use \textbf{clustered standard error} at a cross-sectional level.
\end{frame}

%%%%%%%%%%%
\end{document}
